Testing Gravity using Type Ia Supernovae Discovered by LSST

### 2018-12-18 doc/article.tex
draft for larger group

\title{Testing Gravity using Type Ia Supernovae Discovered by LSST}
\author[0000-0001-6315-8743]{A.~G.~Kim}
\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{S.~BenZvi}
\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
\author{C.~Harper}
\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{D.~Huterer}
\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
Arbor, MI 48109, USA }
\author{C.~Ju}
\affiliation{    Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}

@@

The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the baryonic structures
in the Universe and from the peculiar velocities of test masses therein.
Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
and repulsion of densitiy inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
is determined from its observer magnitude and redshift. For a given a background cosmology, the observed magnitude provides
an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and observer redshifts.

@@

Baryonic structures are sensitive to $f\sigma_8$ through redshift space distortions (RSD), 
which gives the combination $(b + f \mu^2)\sigma_8$ where $b$ is the bias between the tracer and dark matter
and $\mu$ gives the angular separation.  Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and are so bias-free.
As
mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
leads to the dependence on $H_0 d_L$. 
Baryonic structures and peculiar velocities within the same volume are induced by the same overdensities  making their cross-correlation 
insensitive to sample variance \citep{2007PhRvL..99h1301G}: galaxy and peculiar velocity surveys provide synergistic constraints more powerful
than their naive sum.

The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
The current generation of peculiar velocity studies use $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag. 
Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample size by an order of magnitude
over half the sky to a depth of $z=0.1$.

The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
Nevertheless, their low intrinsic-magnitude uncertainties can 
can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
\citep{2008MNRAS.389L..47A,2015JCAP...12..033H, 2017JCAP...05..015H}.
However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.

@@

The precision in  $f\sigma_8$ derived from the baseline WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
from both their RSD and peculiar velocities.
The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance uncertainty is assumed for each supernova.
SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy, cluster, and Ly$\alpha$
RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<-.3$  was well as galaxy surveys will at $z \sim 0.6$ using RSD. 
{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
% Requires the booktabs if the memoir class is not being used
\begin{table}
   \centering
   %\topcaption{Table captions are better up top} % requires the topcapt package
   \begin{tabular}{@{} lccc @{}} % Column formatting, @{} suppresses leading/trailing space
	\hline
	Redshift & RSD & RSD + PV & cumulative\\ \hline
      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
      \hline
   \end{tabular}
   \caption{Projected percent uncertainties in $f\sigma_8$ from a 10-year LSST SN survey with 5\% distance uncertainties from
   \citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
   The cumulative column shows the effective uncertainty from combining the redshift and all other shallower redshift bins.}
   \label{tab:howlett}
\end{table}

@@

\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
\subsection{Survey Parameters that Determine Scientific Performance}
\citet{2017ApJ...847..128H} project how well peculiar velocity surveys can measure $f\sigma_8$. 
They perform their analysis in Fourier space, 
where the Fisher information matrix of a random Gaussian field with mean zero and covariance $C(k)$ parameterized by $\lambda$ is
\begin{equation}
F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right],
\end{equation}
where the covariance for the velocity-velocity correlation
\begin{equation}
C = P_{vv}(k) + \frac{\sigma^2}{n}
\label{cov:eq}
\end{equation}
is dependent on the power spectrum, noise in the velocity measurement dominated by the intrinsic velocity dispersion $\sigma$, and the density of velocity probes
\citep{2017MNRAS.464.2517H}.  
The primary dependence on the growth of structure is through the normalization of the velocity power spectrum, such that 
\begin{equation}
 \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
\end{equation}
for the parameter choice
$\lambda=(f\sigma_8)^2$.  In the sample-variance limit, $F_{\lambda \lambda} \propto V$ whereas in the shot noise limit $F_{\lambda \lambda} \propto V n^2 \sigma^{-4}$.

The survey parameters that  enter the Fisher matrix are the volume $V$, which
in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related by $\sigma_{M} = \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a non-tunable parameter, which only increases with survey duration.
Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.

\begin{itemize}
\item Solid Angle $\Omega$:
The variance in $f\sigma_8$ is proportional to solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole, could double
the sky coverage and halve the variance in $f\sigma_8$.
\item Redshift Depth $z_{max}$:
For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
limited.  We have run simulations of LSST SN surveys
and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
of  Figure~\ref{notshot:fig}, that an increase in the number the number density of tracers leads to a significant improvement in STON.
It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$.

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities.  For several limiting redshift depths, the STON is plotted as a
function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
\label{notshot:fig}}
\end{figure}

@@

\item Intrinsic Magnitude Dispersion $\sigma_M$:
A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$,
as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
limited so the $\sigma^2_M/n$-term
in the Fisher matrix is important.
Survey follow-up stratey,
which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration.

The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively we consider
a pure SN~Ia sample obtained through
spectroscopic classification.  The feasibility of using photometric classification is an important subject of research
that can have implications for survey planning.
\end{itemize}

@@

\begin{itemize}
\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
\item Role of LSST:
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence of 
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
The wide-field of LSST is
 well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
$z<0.3$ SNe~Ia.  LSST is  suited for maximizing $\Omega$ while other (cheaper) resources can improve $\sigma_M$.
\item Follow-up resources to determine per-SN distances:
Supplemental non-LSST follow-up data in the form or improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.

@@

\section{Conclusions}
SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  In the next decade.
high-cadence, wide-field imaging surveys, together with improved precision in their distance determinations, will make SNe~Ia powerful probes
of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $f \sigma_8$ as well
as galaxy surveys, but at lower redshifts that provide better leverage to test gravity models.
While imaging surveys will provide a steady stream of SNe~Ia, a coordinated plan of follow-up is required to take advantage of
their probative power.  The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
are of the highest interest, so that a modest suite of $\sim 5$ 2--4m telescopes should already provide leading measurements of
$f\sigma_8$ at $z\sim 0.1$.

>>>>

\title{Testing Gravity using Type Ia Supernovae Discovered by LSST}
\author[0000-0001-6315-8743]{A.~G.~Kim}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{S.~BenZvi}
\affiliation{Department of Physics and Astronomy, University of Rochester, Rochester, NY 14627, USA}
\author{C.~Harper}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{D.~Huterer}
\affiliation{Department of Physics, University of Michigan, 450 Church Street, Ann
Arbor, MI 48109, USA }
\author{C.~Ju}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}
\author{E.~V.~Linder}
\affiliation{Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720}

@@

In the late 1990's, Type~Ia supernovae (SNe~Ia) were used as distance probes to measure the homogeneous expansion history of the Universe.  The remarkable discovery
that the expansion is accelerating  has called into question our basic understanding of the gravitational forces within the Universe.  Either it
is dominated by a ``dark energy'' that is gravitationally repulsive, or General Relativity is inadequate and needs to be replaced by a modified theory of
gravity.  It is only appropriate that in the upcoming decade, with their sheer numbers and improved distance precisions, SNe~Ia will provide measurements of the {\it inhomogeneous} motions of structures in the Universe
that will provide an unmatched test of whether dark energy or modified gravity is responsible for the accelerating expansion of the Universe.

The growth of structure depends on the expansion history of the Universe, the nature and density of its contents, and gravity. 
It is therefore a powerful probe of cosmology and dark energy.  Growth of structure can be measured from the light-emittting structures
in the Universe and from the peculiar velocities of test masses therein.
Peculiar velocities are the motions, on top of the cosmological expansion, caused by the gravitational attraction
and repulsion of density inhomogeneities in the Universe.  The peculiar velocity of an object with a known absolute magnitude
can be determined from its observer-frame magnitude and redshift. For a given background cosmology, the observed magnitude provides
an estimate of the cosmological redshift, the peculiar velocity is then the difference between the cosmological and CMB-frame redshifts.

@@

Baryonic structures are sensitive to $f\sigma_8$ through redshift space distortions (RSD), 
which gives the combination $(b + f \mu^2)\sigma_8$ where $b$ is the bias between the tracer and dark matter
and $\mu$ gives the angular separation of Fourier modes to the line-of-sight.  Correlations between peculiar velocities are sensitive to  $f\sigma_8 (H_0 d_L)^{-1}$.
Although the velocities may be measured off of biased tracers, the tracers' dynamics are driven by all mass (including dark matter) and so are bias-free.
As
mentioned earlier, peculiar velocities are measured relative to the background cosmological expansion that
leads to the dependence on $H_0 d_L$. 
Baryonic structures and peculiar velocities within the same volume are induced by the same overdensities  making their cross-correlation 
insensitive to sample variance \citep{2007PhRvL..99h1301G}: galaxy and peculiar velocity surveys provide synergistic constraints more powerful
than their naive sum.

The probative power of a specific peculiar-velocity tracer primarily depends on its number density and the precision to which its absolute magnitudes are known.
The current generation of peculiar velocity studies uses $10^3-10^5$ galaxies with Fundamental Plane and Tully-Fisher distances \citep{2008AJ....135.1738M, 2014MNRAS.445.2677S,
2016AJ....152...50T}.  These galaxies have absolute magnitude uncertainties of $\sim 0.4$~mag. 
Next generation surveys WALLABY \citep{2008ExA....22..151J} and TAIPAN \citep{2017PASA...34...47D} are designed to increase these sample sizes by an order of magnitude
over half the sky to a depth of $z=0.1$.

The current sample of SNe~Ia has a low number density compared to Fundamental Plane and Tully-Fisher galaxies.
Nevertheless, their low intrinsic-magnitude uncertainties 
can provide peculiar velocities (expressed equivalently as peculiar magnitudes)
of their host galaxies \citep{2006PhRvD..73l3526H,2011ApJ...741...67D}.  Existing SN~Ia samples
have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
\citep{2008MNRAS.389L..47A,2015JCAP...12..033H, 2017JCAP...05..015H}.
However, the signal-to-noise is currently insufficient to perform a meaningful test of GR.

@@

The precision in  $f\sigma_8$ derived from the baseline WFD ten-year SN~Ia discoveries has been projected by \citet{2017ApJ...847..128H},
from both their RSD and peculiar velocities.
The results are summarized in Table~\ref{tab:howlett}, where a 5\% distance (0.1~mag) uncertainty is assumed for each supernova.
SNe~Ia alone can provide a $2\%$ measurement of $f\sigma_8$ at $z<0.3$ redshifts lower than where galaxy, cluster (through the
kinematic S-Z effect), and Ly$\alpha$
RSD measurements are sensitive.  A joint galaxy RSD and SN peculiar velocity will be even more powerful.
SNe~Ia peculiar velocities will measure $f\sigma_8$ at $z<0.3$  as well as galaxy surveys will at $z \sim 0.6$ using RSD. 
This can be compared to the current 15\% uncertainty from 6dF \citep{2017MNRAS.471..839A}, 3\% uncertainty projected
for the combined TAIPAN and WALLABY+WNSHS surveys \citep{2017MNRAS.464.2517H} for low-redshift measurements of $f\sigma_8$.
At larger $z>0.3$ redshifts DESI projects 10\% accuracies.

%{\bf Add a SN PV + DESI-like BGS column and/or a PV-only column with Cullan's code? }
% Requires the booktabs if the memoir class is not being used
\begin{table}
   \centering
   %\topcaption{Table captions are better up top} % requires the topcapt package
   \begin{tabular}{@{} lccc @{}} % Column formatting, @{} suppresses leading/trailing space
	\hline
	Redshift & RSD & RSD + PV & cumulative\\ \hline
      $0.00<z<0.05$   & 66.3 & 13.9 & 13.9\\
     $0.05<z<0.10$            & 24.6     &  7.3 & 6.5\\
     $0.10<z<0.15$      & 14.8  & 5.8 & 4.3\\
     $0.15<z<0.20$      & 10.6  & 5.0 & 3.3\\
      $0.20<z<0.25$     & 8.3  & 4.4 & 2.6\\
     $0.25<z<0.30$  & 6.8  &  4.0 & 2.2\\
      \hline
   \end{tabular}
   \caption{Projected percent uncertainties in $f\sigma_8$ from a 10-year LSST SN survey with 5\% distance uncertainties from
   \citet{2017ApJ...847..128H}. RSD is from clustering, and RSD~+~PV is from joint clustering and peculiar velocities.
   The cumulative column shows the effective uncertainty from combining its and all other shallower redshift bins.}
   \label{tab:howlett}
\end{table}

@@

\section{Designing a Supernova Survey to Measure $\MakeLowercase{f}\sigma_8$}
\subsection{Survey Parameters that Determine Scientific Performance}
\citet{2017ApJ...847..128H} project how well peculiar velocity surveys can measure $f\sigma_8$. 
They perform their analysis in Fourier space, 
where the Fisher information matrix of parameters $\lambda_i$ from random Gaussian field with mean zero and covariance $C(k)$ is
\begin{equation}
F_{ij} = \frac{V}{2}\int \frac{d^3k}{(2\pi)^3} \text{Tr}\left[ C^{-1} \frac{\partial C}{\partial \lambda_i} C^{-1}
\frac{\partial C}{\partial \lambda_j} \right],
\end{equation}
where the covariance for the velocity-velocity correlation
\begin{equation}
C = P_{vv}(k) + \frac{\sigma^2}{n}
\label{cov:eq}
\end{equation}
is dependent on the power spectrum, noise in the velocity measurement dominated by the intrinsic velocity dispersion $\sigma$, and the density of velocity probes
\citep{2017MNRAS.464.2517H}.  
The primary dependence on the growth of structure is through the normalization of the velocity power spectrum, such that 
\begin{equation}
 \frac{\partial P_{vv}}{\partial \lambda} = \text{constant}
\end{equation}
for the parameter choice
$\lambda=(f\sigma_8)^2$.  In the sample-variance limit, $F_{\lambda \lambda} \propto V$ whereas in the shot noise limit $F_{\lambda \lambda} \propto V n^2 \sigma^{-4}$.

The survey parameters that  enter the Fisher matrix are the volume $V$, which
in turn can be parameterized by the survey solid angle $\Omega$ and the redshift depth $z_{max}$ ; the number density of sources $n$; and the intrinsic
magnitude dispersion $\sigma_M$, where magnitude and velocity dispersions are related at low-redshift by $\sigma_{M} \approx \frac{5}{\ln{10}} \frac{1+z}{z} \sigma$.
LSST is expected to discover all $z<0.3$ SNe~Ia before maximum light in its active Wide Fast Deep (WFD) survey area.
Equating LSST discovery with having a meaningful peculiar velocity, the number density $n$ is thus a fixed parameter, which only increases with survey duration.
Thus $\Omega$, $z_{max}$, and $\sigma_M$ are left as the relevant parameters.

\begin{itemize}
\item Solid Angle $\Omega$:
The variance in $f\sigma_8$ is inversely proportional to solid angle $\Omega$.  The baseline
LSST survey covers 18,000 sq.~deg, at $-75 \lesssim \delta \lesssim15$ avoiding the Galactic plane.  Larger solid-angle coverage for $z<0.3$ SN~Ia searches, beyond
the LSST baseline, benefits peculiar-velocity science.  Complementary northern-hemisphere surveys and LSST-expanded or independent
coverage of the southern equatorial pole, could double
the sky coverage and halve the variance in $f\sigma_8$.
\item Redshift Depth $z_{max}$:
For the number densities generated by LSST and the SN~Ia intrinsic magnitude dispersion, the noise is not sample-variance
limited.  We have run simulations of LSST SN surveys
and calculated the signal-to-noise (STON) of $f\sigma_8$ from the peculiar-velocity correlations. These simulations show, as seen in the left plot
of  Figure~\ref{notshot:fig}, that an increase in the number density of tracers leads to a significant improvement in STON.
It is also worth noting that sample variance is not negligible, as seen by the non-linear dependence of the  STON on $\sigma_M^2/n$.

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth]{../outcosmo/notshot.png}
\includegraphics[width=0.48\textwidth]{../outcosmo/perSN.png}
\caption{Left: STON of $f\sigma_8$ from simulated LSST SN peculiar-velocities only (no RSD).  For several limiting redshift depths, the STON is plotted as a
function of $\sigma_M^2/n$, normalized to intrinsic magnitude dispersion $\sigma_M=0.1$~mag and 10-year LSST number densities.
Right: The effective contribution to the STON per supernova, quantified as STON divided by the root of the number of tracers.
\label{notshot:fig}}
\end{figure}

@@

\item Intrinsic Magnitude Dispersion $\sigma_M$:
A decrease in the supernova intrinsic magnitude dispersion leads to a significant improvement in  the STON of $f\sigma_8$,
as shown in Figure~\ref{notshot:fig} in parallel to our earlier discussion of the dependence of $n$.  We are not sample-variance
limited so the $\sigma^2_M/n$-term
in the Fisher matrix is important.
Survey follow-up strategy,
which determines which data are available per SN~Ia, influences the accuracy to which their distances can be determined. 
This dependence is non-trivial: the improvement between $\sigma_M=0.08$ and 0.15~mag dispersions is equivalent to a factor of 3.52
 in number density, or equivalently in survey duration.

The intrinsic magnitude dispersion depends on the purity of the sample. Conservatively we consider
a pure SN~Ia sample obtained through
spectroscopic classification, as the requisite follow-up resources for a  complete $z<0.3$ sample is within reason. 
The feasibility of using photometric classification is an important subject of research
that can have implications for survey planning, particularly in extending the redshift range.
\end{itemize}

@@

\begin{itemize}
\item Spectroscopic transient classification:  SNe~Ia are defined through the absence of Hydrogen and the presence of the 6150~\AA~Silicon P-Cygni feature.
\item Spectroscopic redshifts:  Peculiar velocities are extracted directly from redshift measurements.  Redshift uncertainties of $>0.5$\% contribute significantly
to the error budget.  We thus conclude that there is the need for spectroscopic $R>200$ host-galaxy redshifts.  In most cases this redshift will be
available in the classification spectrum.  Planned bright-galaxy redshift surveys can also observe a significant fraction of nearby SN hosts.
\item Role of LSST:
LSST can be taken to be a SN~Ia discovery and distance machine, or a discovery machine only.   The nominal cadence  
 produces sparse per-band light curves, which limits the intrinsic magnitude dispersion possible with LSST data alone to $\sigma_M \approx 0.15$~mag.
The wide-field of LSST is
 well suited for discoveries over a larger area of sky, but does not provide significant multiplex advantage for the low surface-density of active
$z<0.3$ SNe~Ia.  LSST is  suited for maximizing $\Omega$ while other (cheaper) resources can improve $\sigma_M$.
\item Follow-up resources to determine per-SN distances:
Supplemental non-LSST follow-up data in the form of improved temporal light-curve sampling (accurate rise and decline times), 
expanded (UV, NIR) wavelength coverage, and spectral features
can access high-fidelity SN~Ia models with lower intrinsic magnitude dispersion and residual systematic bias.
For example, infrared data \citep{2012MNRAS.425.1007B} or spectrophotometry at peak brightness
 \citep{2015ApJ...815...58F} are projected to give $\sigma_M \lesssim 0.08$~mag.

@@

\section{Conclusions}
SNe~Ia are already powerful probes of the homogeneous cosmological expansion of the Universe.  In the next decade.
high-cadence, wide-field imaging surveys, together with improved precision in their distance determinations, will make SNe~Ia powerful probes
of the gravity induced-motion caused by the inhomogeneous Universe.  SNe~Ia peculiar velocities at $z<0.3$ will measure $f \sigma_8$ significantly better
than galaxy surveys, at lower redshifts that provide better leverage to test gravity models.
While imaging surveys will provide a steady stream of SNe~Ia, a coordinated plan of follow-up is required to take advantage of
their probative power.  The resources necessary to follow hundreds of thousands of SNe depend on specific follow-up choices, and 
access to those resources will define the redshift limits of the survey.  Fortunately, the lowest-redshift, and hence brightest, supernovae
are of the highest interest, so that a modest suite of $\sim 5$ 2--4m telescopes should be capable of providing leading measurements of
$f\sigma_8$ at $z\sim 0.1$.
